Template Master's Thesis

### 2023-08-12 paper.tex
Update introduction and data collection

Public permissionless blockchains such as Bitcoin \citep{nakamotoBitcoin2008} and Ethereum \citep{buterin2014ethereum} allow users %individuals
 to participate with multiple pseudonymous addresses. The creation of these addresses is virtually cost free. Contrary to popular belief, these blockchains are entirely transparent. All transactions are stored as part of the blockchain's history and publicly observable.
This has opened up a nascent scientific field dealing with entity identification within blockchain networks. Researchers try to cluster addresses controlled by the same user by analyzing on-chain data and detecting usage patterns. The frequent lack of ground truth makes it difficult to evaluate different clustering methods. As a result, most methods are heuristic. %The majority of these methods are heuristic which often makes them difficult to evaluate due to the absence of any ground truth.
\newline On one hand, identifying the addresses that belong to the same real-world entity is beneficial. According to \cite{FV:17} it allows better evaluation of network properties with respect to usage, distribution of wealth and detecting fraudulent activities. For instance, if a user distributes their voting power across various addresses, they might manipulate an on-chain voting process that seems fair at the outset. \newline
Conversely, the lack of privacy is also detrimental to most financial use cases (DeFi). As a response, multiple privacy-enhancing tools and protocols have been proposed to obfuscate transaction tracing. 
Nonetheless, these protocols are not yet widely adopted and careless user behaviour can undermine the privacy guarantees they offer. \newline
%As such/In conclusion
The topic of privacy, or lack thereof, will remain significant for future blockchain development and research.

%The most prominent/widely used non-custodial mixer is Tornado cash (Schär, Tutela).
%If someone obtains information that allows them to link a blockchain address to an entity, they may effectively observe that entity’s entire transaction history and associated activity. Even if the entity uses multiple addresses, any link between these addresses may expose the fact that they belong to the same person.

\textbf{Related work}\\
Previous work in this field can be broadly distinguished between methods %heuristics
 developed for the unspent transaction output (UTXO) model (e.g, Bitcoin and ZCash) and the account-based model (e.g., Ethereum and Polygon). While both models share the concept of addresses, the notion of accounts is not present in UTXO-based blockchains. The way in which transactions are processed is fundamentally different, and therefore, clustering heuristics are not applicable to both paradigms.

A number of address clustering heuristics in the UTXO setup have been proposed for Bitcoin and derivatives \citep{Androulaki2013, Meiklejohn2013, Haslhofer2016, jourdan2018, kappos2022}. All these methods will not be the focus of this work and therefore not discussed. \newline 
As suggested by \cite{nakamotoBitcoin2008} in the Bitcoin whitepaper, most Bitcoin wallet implementations use a new key pair for each transaction to keep them from being linked to a common owner. In contrast to the UTXO-model, native transactions in account-based blockchains can only move funds between a single sender and receiver, and the ``change`` remains in the sender account. Subsequent transactions necessarily use the same address again. The account-based model essentially relies on address reuse on the protocol level \citep{Beres2020}. Therefore, privacy guarantees should be lower as most participants are likely to use only a small number of addresses. \newline
% Make an assumption that people only use a small set of addresses
Clustering heuristics for account-based blockchains were first introduced by \cite{FV:17}. He proposed and applied heuristics that exploit patterns related to deposit addresses, multiple participation in airdrops, and token authorization mechanisms. \newline
\cite{Beres2020} propose more universally applicable methods, as they argue that Victor’s heuristics, while powerful, assume participation in certain on-chain events. Their technique interprets transactions or token transfers as network graphs, with addresses as nodes and asset transfers as edges. They quantitavely compare graph-representation learning algorithms (a subset of machine learning) and propose further user profiling techniques based on time-of-day activity and transaction fees. Using ENS address pairs as ground truth, the authors rigorously test their methods and apply their findings to significantly reduce the privacy guarantees of \textit{Tornado Cash}, a non-custodial privacy-enhancing protocol on Ethereum. \newline
\cite{wu2022tutela} extend on one of \cite{Beres2020} graph representation learning %node-embedding
 algorithms and apply it at scale. Further, they also propose a set of new clustering heuristics targeting Tornado Cash,  showing that careless user behavior can still reveal identity. Based on those heuristics, they built an application to measure the anonymity of an Ethereum address. \newline All of the methods mentioned will be discussed in greater detail in Section 3. \newline
Broadening the scope of entity identification, \cite{victorlüders2019} study the largest Ethereum ERC-20 token networks from a graph perspective. Similarly, \cite{casalebrunet2021} apply network graph analysis to various non-fungible token (NFT) ecosystems. Both find that many of the networks follow either a star or a hub-and-spoke pattern. These patterns are common to interaction graphs in social networks. Further, \cite{Payette2017} propose a segmentation of the Ethereum address space into four distinct behaviour groups sharing similar attributes while using k-means clustering. \newline
Rather than treating users as entities and analyzing on-chain data, \cite{yu2023} propose (a novel) approach for correlation analysis by exploiting network information. Although this approach has the potential to avoid the impact of privacy-enhancing technologies, it introduces a new set of limitations and problems. Thus, this approach may be of great interest for future research, particularly when privacy-enhancing techniques become more widely used.

@@

\subsection{Decentraland}
\textit{Decentraland} is the first large-scale blockchain-based virtual world. Well-known companies like Tommy Hilfiger, Samsung, PepsiCo, Diesel, Adidas, and Netflix, who are actively engaging in Decentraland  \citep{metaverse-retailing2023}.\newline
Decentraland is an ideal platform for empirical research for two main reasons.
\textit{First}, its open architecture allows compiling snapshots of user activity, including the users' locations  within the metaverse. This was done by \cite{metaverse-retailing2023}, who captured snapshots of avatar locations over a period of 10 months. The addresses they collected are used for this thesis.
\textit{Second}, users who connect their Web3 account can own various digital assets, such as monetary units, land parcels, and avatar collectibles (e.g. wearables, emotes, and names). Smart contracts on Ethereum and Polygon track and manage the ownership of these assets \citep{goldbergschaer2023}. This allows us to analyze the entire transaction history and derive information about the users' economic activity. %avatars'
\newline
Initially, Decentraland launched with two native tokens. \textit{LAND}, an NFT compliant with ERC-721, manages the ownership of land parcels in Decentraland. \textit{MANA}, a fungible token compliant with ERC-20, is the in-world currency used to purchase digital goods and services. The majority of purchases and trades are settled in MANA. \newline 
Moreover, the Decentraland core developers provide a smart-contract-based marketplace where users may buy or exchange LAND or other in-game collectibles. %developers/core team

@@

\subsection{Non-custodial crypto asset mixers}
According to \cite{nadler2023tornado}, crypto-asset mixers are currently the most widely used approach to achieve privacy on public blockchains. At a high level, we can distinguish between custodial and non-custodial crypto-asset mixers. \newline
In the custodial setup, users send their assets to a centralized service provider's public deposit address. After some delay, the provider sends the assets back to a privately relayed recipient address. This approach is fully trust-based, as the service provider has complete control over the assets and access to the identifying data. \newline 
In contrast, non-custodial crypto asset mixers (e.g., Tornado Cash) replace the trusted mixing party with a publicly verifiable smart contract. They rely on cryptographic schemes that allow anyone to prove and independently verify the validity of a withdrawal without disclosing the link to a specific deposit. This has the advantage that users do not have to share the identifying information with anyone and there is no liquidity risk as the funds are locked. \citep{nadler2023tornado} \newline
In a nutshell, the mixer works in the following way: Various entities deposit the same amount of a specific crpyto asset to a mixer address acting as a pool. Anyone who has contributed to the pool may then generate a new address and withdraw their funds without revealing the link between the deposit and withdrawal addresses. This is achieved using zkSNARKs (zero knowledge, succinct, non-interactive argument of knowledge). 
Each depositor inserts a hash value in a Merkle-tree. At the time of withdrawal, each legitimate withdrawer can prove unlinkably with a zero-knowledge proof that they know the pre-image of a previously inserted hash leaf in the Merkle-tree. %Subsequently, users can withdraw their asset from the mixer whenever they consider that the size of the anonymity set is satisfactory.
 For a more detailed description of how Tornado Cash works, see \cite{nadler2023tornado} or \cite{Beres2020}. \newline
In the context of this work, it is sufficient to know that third parties can still observe the addresses that have deposited to and withdrawn from the pool. The set of all deposits that a particular withdrawal could be originated from is defined as the \textit{anonymity set}. 
In theory, a larger anonymity set will provide a higher level of privacy, since  any withdrawal address could be linked to more deposit addresses. However, in practice this does not always hold true. %this is not always the case
Some users reveal the link between their withdrawal and deposit addresses, which also affects the privacy of other depositors. For example, by withdrawing from the pool with the same address they deposited or directly send funds between the two addresses. According to \cite{nadler2023tornado}, careless use of the protocol, testing transactions or external incentive mechanisms might be reasons for this behaviour. \newline
In addition, \cite{Beres2020} and \cite{wu2022tutela} propose approaches to de-anonymize Tornado cash transactions using various clustering heuristics. These methods are discussed in more detail in section 4.

@@

\section{Data Collection and Preparation}
% Introduction Satz - Ausgangslage 
The starting point for our work was a dataset comprising 473,927 distinct (Web3) addresses. %The original dataset comprised 473,927 distinct (Web3) addresses.
 As we focus on clustering methods using on-chain data, we exclude addresses that have not been recorded on either Ethereum or Polygon. To accomplish this, we used data from Blockscan\footnote{\url{https://blockscan.com/}}. Overall, 59,651 addresses were recorded on Ethereum, 129,988 on Polygon, with 52,095 addresses appearing on both networks simultaneously. The set of ``clusterable`` addresses therefore counts 137,544 addresses. Subsets are visualized as a Venn diagram in Figure \ref{fig:Venn}.

@@

\subsection{Transactions and Token Transfer Events}
The address subsets were used to collect transaction and token transfer event data from Ethereum and Polygon. Ethereum or Polygon data can be accessed directly through a node or an application programming interface (API) provider like Infura or Etherscan. The absence of account indexing in Ethereum or Polygon poses a challenge for retrieving all past transactions and token transfers of a specific address, as it requires scanning through all blocks and token transfer events emitted by designated token contracts. Fortunately, Etherscan offers an API Endpoint Module for ``Accounts`` that facilitates the retrieval of transactions and token transfer events for a given address. By using our network-specific address subsets, we were able to significantly decrease the number of required API calls. 
We gathered all transactions and token transfers up until block \texttt{17,670,000} on Ethereum (July 11, 2023) and block \texttt{44,990,000} on Polygon (July 12, 2023). In total, we collected more than 30 million %30,689,978
token transfer events and more than 16 million %16,092,531
 normal transactions. While the amount of transactions is similar between the two networks, there are three times more transfer events on Polygon. \newline The output was stored in separate CSV files and also imported into a MongoDB\footnote{\url{https://www.mongodb.com/}} database. The database contains distinct collections for transactions and transfers. For additional information on the different data fields, see Appendix.

@@

\subsection{Intra-set asset transfers}
In this project, we want to use entity clustering to determine the number of entities that are represented by the remaining 137,544 addresses. Or, to phrase it differently, do users interact with Decentraland using multiple EOAs, and can we identify addresses that belong to the same user?
\newline
To achieve this, we often only need to consider transactions or token transfer events where both the ``\texttt{from}`` (sender) and the ``\texttt{to}`` (recipient) address are within the address set. Regarding normal transactions, this means that we only keep native asset transfers (Ether or Matic) as almost all addresses are EOAs. \newline
The intra-set asset transfer dataset is mainly used to generate the network graphs (see section xx). When clustering addresses using graph-based approaches, this has the advantage that we do not need to differentiate between EOAs and smart contract account later on. Furthermore, it allows us to observe and visualize the asset flows between users. In total, we have 974,478 intra-set token transfers and 210,548 intra-set native asset transfers (101,621 Matic, 108,927 Ether).

\iffalse
%As nearly all addresses are Externally Owned Accounts (EOAs), we do not need to differentiate between EOAs and smart contract accounts, which further simplifies the clustering process. \newline
%(logging in to Decentraland with multiple addresses) When it comes to entity clustering, we essentially want to answer the question of how many real-world entities our remaining  137,544 addresses represent. 
-Also for other clustering heuristics we often make this reduction, for example: In the self-approval approach, we only consider approval transactions where another address from the set is approved.
-filter our transactions and token transfers to entries where both the 'from' and the 'to' are within this narrowed pool of addresses. We define this as intra-set asset transfers. For the transactions, this means that we only consider native asset (Ether or matic) transfers. %between two EOAs
-As already mentioned in the introduction, the goal of this paper is to cluster addresses within who (1) were active in Decentraland during the 9-month period (e.g., within the initial dataset) and (2) transacted or interacted with any token on Ethereum or Polygon. 
-Essentialy, we want to answer two main research questions when it comes to clustering:
1. How many unique entities does the address set represent (i.e. were active on Decentraland during the 9-month period)?
2. Are users interacting through multiple accounts/EOAs/addresses with Decentraland?
-we focus on on-chain data (and not user location or avatar data), allowing us to exclude addresses that have not transacted or interacted with any token on Ethereum or Polygon. 
-As mentioned in the introduction, the goal of this paper is to cluster addresses belongig to the same entity that were recorded/active in Decentraland during the 9-month period. Therefore, we disregard addresses outside of this set. Also, since we only look at on-chain data, we cannot cluster addresses without activity on Polygon or Ethereum. Doing these two reductions, there are  141, 591 'clusterable' addresses remaining.
-Since all of our addresses are EOAs (some multisig smart contract wallets). -> Check this?
\fi

@@

Figure \ref{fig:ENS} illustrates the mechanism that allowed us to generate ground truth pairs in a simplified way: Initially, the domain name \texttt{`foo.eth`} resolves to address \texttt{`0xABC...`}. The owner has also set a reverse resolution for this address, indicated by the the edge from the address to the domain name. Next, the owner (he) decides to change the resolved address to \texttt{`0xDEF...`}. Again, he is setting a reverse record.  As a domain can only point to one address at the time, the previous mapping from address to old address is deleted. However, \texttt{`ABC....addr.reverse`} still points to \texttt{`foo.eth`} and when we query the name corresponding to \texttt{`0xABC...`} and \texttt{`0xDEF...`}, both will return \texttt{`foo.eth`}.
If this special case occurs, we assume that both addresses belong to the same entity.
%We did not observe cases where someone set a reverse record to a completely unrelated address (e.g. alice.eth)

@@

We applied the following methodology to find ground truth pairs: First, we collected all addresses that interacted ENS (NFTs) to reduce the amount of name look-ups. For each of the remaining addresses, we checked if it points to an ENS domain name\footnote{This was done via the ENS module within web3.py \url{https://web3py.readthedocs.io/en/stable/ens_overview.html}}. If it did, we stored the associated ENS domain alongside the address in a CSV file. %(It provides an interface to look up domains and addresses). 
Using the newly generated list, we checked if two addresses point to the same ENS domain. If this check was met, we stored both addresses alongside the ENS domain. \newline
We found 11,440 reverse records and 40 address pairs. %(addresses pointing to the same .eth name)
We use these pairs as our ground truth to measure the performance of different clustering heuristics. Even though this is a small sample compared to the total number of clusterable addresses, we should still be able to evaluate (compare) the methods and infer their efficacy. %profiling techniques

>>>>

Public permissionless blockchains such as Bitcoin \citep{nakamotoBitcoin2008} and Ethereum \citep{buterin2014ethereum} allow users %individuals
 to participate with multiple pseudonymous addresses, the creation of which is virtually cost free. Contrary to popular belief, these blockchains are entirely transparent. All transactions are publicly observable and stored as part of the blockchain's history.
This has opened up a nascent scientific field dealing with entity identification within blockchain networks. Researchers try to cluster addresses controlled by the same user by analyzing on-chain data and detecting usage patterns. The frequent lack of ground truth makes it difficult to evaluate different clustering methods. As a result, most methods are heuristic. %The majority of these methods are heuristic which often makes them difficult to evaluate due to the absence of any ground truth.
\newline On one hand, identifying the addresses that belong to the same real-world entity is beneficial. According to \cite{FV:17} it allows better evaluation of network properties with respect to usage, distribution of wealth and detecting fraudulent activities. For instance, if a user distributes their voting power across various addresses, they might manipulate an on-chain voting process that seems fair at the outset. \newline
Conversely, the lack of privacy is also detrimental to most financial use cases (DeFi). As a response, multiple privacy-enhancing tools and protocols have been proposed to obfuscate transaction tracing. 
Nonetheless, these protocols are not yet widely adopted and careless user behaviour can undermine the privacy guarantees they offer. \newline
With these considerations in mind, the topic of privacy, or lack thereof, will remain significant for future blockchain development and research.

%If someone obtains information that allows them to link a blockchain address to an entity, they may effectively observe that entity’s entire transaction history and associated activity. Even if the entity uses multiple addresses, any link between these addresses may expose the fact that they belong to the same person.

\textbf{Related work}\\
Previous work in this field can be broadly distinguished between methods developed for the unspent transaction output (UTXO) model (e.g, Bitcoin and ZCash) and the account-based model (e.g., Ethereum and Polygon). While both models share the concept of addresses, the notion of accounts is not present in UTXO-based blockchains. The way in which transactions are processed is fundamentally different, and therefore, clustering heuristics are not applicable to both paradigms.

Several address clustering heuristics in the UTXO setup have been proposed for Bitcoin and its derivatives \citep{Androulaki2013, Meiklejohn2013, Haslhofer2016, jourdan2018, kappos2022}. However, these methods are outside the scope of this work and will not discussed. \newline 
As suggested by \cite{nakamotoBitcoin2008} in the Bitcoin whitepaper, most Bitcoin wallet implementations use a new key pair for each transaction to keep them from being linked to a common owner. Unlike the UTXO-model, native transactions in account-based blockchains can only move funds between a single sender and receiver, and the ``change`` remains in the sender account. Subsequent transactions necessarily use the same address again. The account-based model essentially relies on address reuse on the protocol level \citep{Beres2020}. Consequently, privacy guarantees should be lower as participants typically use a limited set of addresses. \newline
% Make an assumption that people only use a small set of addresses
Clustering heuristics for account-based blockchains were first introduced by \cite{FV:17}. He proposed and applied heuristics that exploit patterns related to deposit addresses, multiple participation in airdrops, and token authorization mechanisms. \newline
\cite{Beres2020} propose more universally applicable methods, arguing that Victor’s heuristics, while powerful, assume participation in certain on-chain events. Their approach interprets transactions or token transfers as network graphs, with addresses as nodes and asset transfers as edges. They quantitavely compare graph-representation learning algorithms (a subset of machine learning) and propose further user profiling techniques based on time-of-day activity and transaction fee patterns. Using ENS address pairs as ground truth, the authors rigorously test their methods and apply their findings to significantly reduce the privacy guarantees of \textit{Tornado Cash}, a non-custodial privacy-enhancing protocol on Ethereum. \newline
\cite{wu2022tutela} extend on one of \cite{Beres2020} graph representation learning %node-embedding
 algorithms and apply it at scale. Further, they also propose a set of new clustering heuristics targeting Tornado Cash,  showing that careless user behavior can still reveal identity. Based on those heuristics, they developed an application to measure the anonymity of an Ethereum address. \newline All of the aforementioned methods will be discussed in greater depth in Section 3. \newline
Broadening the scope of entity identification, \cite{victorlüders2019} study the largest Ethereum ERC-20 token networks from a graph perspective. Similarly, \cite{casalebrunet2021} apply network graph analysis to various non-fungible token (NFT) ecosystems. Both find that many of the networks follow either a star or a hub-and-spoke pattern, common to interaction graphs in social networks. Moreover, \cite{Payette2017} propose a segmentation of the Ethereum address space into four distinct behaviour groups sharing similar attributes while using k-means clustering. \newline
Rather than treating users as entities and analyzing on-chain data, \cite{yu2023} propose (a novel) approach for correlation analysis by exploiting network information. Although this approach has the potential to avoid the impact of privacy-enhancing technologies, it introduces a new set of limitations and problems. Thus, this approach may be of great interest for future research, particularly when privacy-enhancing techniques become more widely used.

@@

\subsection{Decentraland}
\textit{Decentraland} is the first large-scale blockchain-based virtual world. Well-known companies like Tommy Hilfiger, Samsung, PepsiCo, Diesel, Adidas, and Netflix, are actively engaging in Decentraland  \citep{metaverse-retailing2023}.\newline
Decentraland is an ideal platform for empirical research for two main reasons.
\textit{First}, its open architecture allows compiling snapshots of user activity, including the users' locations  within the metaverse. This was done by \cite{metaverse-retailing2023}, who captured snapshots of avatar locations over a period of nine months. The addresses they collected are used for this thesis.
\textit{Second}, users who connect their Web3 account can own various digital assets, such as monetary units, land parcels, and avatar collectibles (e.g. wearables, emotes, and names). Smart contracts on Ethereum and Polygon track and manage the ownership of these assets \citep{goldbergschaer2023}. This allows us to analyze the entire transaction history and derive information about the users' economic activity. %avatars'
\newline
Initially, Decentraland launched with two native tokens. \textit{LAND}, an NFT compliant with ERC-721, manages the ownership of land parcels in Decentraland. \textit{MANA}, a fungible token compliant with ERC-20, is the in-world currency used to purchase digital goods and services. The majority of purchases and trades are settled in MANA. \newline 
Moreover, the Decentraland core developers provide a smart-contract-based marketplace where users may buy or exchange LAND or other in-game collectibles. %developers/core team

@@

\subsection{Non-custodial crypto asset mixers}
According to \cite{nadler2023tornado}, crypto-asset mixers are currently the most widely used approach to achieve privacy on public blockchains. At a high level, we can distinguish between custodial and non-custodial crypto-asset mixers. \newline
In the custodial setup, users send their assets to a centralized service provider's public deposit address. After some delay, the provider sends the assets back to a privately relayed recipient address. This approach is fully trust-based, as the service provider has complete control over the assets and access to the identifying data. \newline 
In contrast, non-custodial crypto asset mixers (e.g., Tornado Cash) replace the trusted mixing party with a publicly verifiable smart contract. They rely on cryptographic schemes that allow anyone to prove and independently verify the validity of a withdrawal without disclosing the link to a specific deposit. This has the advantage that users do not have to share the identifying information with anyone and there is no liquidity risk as the funds are locked. \citep{nadler2023tornado} \newline
In a nutshell, the mixer works in the following way: Various entities deposit the same amount of a specific crpyto asset to a mixer address acting as a pool. Anyone who has contributed to the pool may then generate a new address and withdraw their funds without revealing the link between the deposit and withdrawal addresses. This is achieved using zkSNARKs (zero knowledge, succinct, non-interactive argument of knowledge). 
Each depositor inserts a hash value in a Merkle-tree. At the time of withdrawal, each legitimate withdrawer can prove unlinkably with a zero-knowledge proof that they know the pre-image of a previously inserted hash leaf in the Merkle-tree. %Subsequently, users can withdraw their asset from the mixer whenever they consider that the size of the anonymity set is satisfactory.
 For a more detailed description of how Tornado Cash works, see \cite{nadler2023tornado} or \cite{Beres2020}. \newline
In the context of this work, it is sufficient to know that third parties can still observe the addresses that have deposited to and withdrawn from the pool. The set of all deposits that a particular withdrawal could be originated from is defined as the \textit{anonymity set}. 
In theory, a larger anonymity set will provide a higher level of privacy, since  any withdrawal address could be linked to more deposit addresses. However, in practice this does not always hold true. %this is not always the case
Some users reveal the link between their withdrawal and deposit addresses, which also affects the privacy of other depositors. For example, by withdrawing from the pool with the same address they deposited or directly sending funds between the two addresses. According to \cite{nadler2023tornado}, careless use of the protocol, testing transactions or external incentive mechanisms might be reasons for this behaviour. \newline
In addition, \cite{Beres2020} and \cite{wu2022tutela} propose approaches to de-anonymize Tornado cash transactions using various clustering heuristics. These methods are discussed in more detail in section 4.

@@

\section{Data Collection and Preparation}
% Introduction Satz - Ausgangslage 
The starting point for our work was a dataset comprising 473,927 distinct (Web3) addresses. %The original dataset comprised 473,927 distinct (Web3) addresses.
 As we focus on clustering methods using on-chain data, we exclude addresses that have not been recorded on either Ethereum or Polygon. To accomplish this, we used data from Blockscan\footnote{\url{https://blockscan.com/}}. Overall, 59,651 addresses were recorded on Ethereum, 129,988 on Polygon, with 52,095 addresses appearing on both networks simultaneously. The set of ``clusterable`` addresses therefore counts 137,544 addresses. The subsets are visualized as a Venn diagram in Figure \ref{fig:Venn}.

@@

\subsection{Transactions and Token Transfer Events}
The address subsets were used to collect transaction and token transfer event data from Ethereum and Polygon. Ethereum or Polygon data can be accessed directly through a node or an application programming interface (API) provider like Infura or Etherscan. The absence of account indexing in Ethereum or Polygon poses a challenge for retrieving all past transactions and token transfers of a specific address, as it requires scanning through all blocks and token transfer events emitted by designated token contracts. Fortunately, Etherscan offers an API Endpoint Module for ``Accounts`` that facilitates the retrieval of transactions and token transfer events for a given address. By using our network-specific address subsets, we were able to significantly decrease the number of required API calls. 
We gathered all transactions and token transfers up until block \texttt{17,670,000} on Ethereum (July 11, 2023) and block \texttt{44,990,000} on Polygon (July 12, 2023). In total, we collected more than 30 million %30,689,978
token transfer events and more than 16 million %16,092,531
 normal transactions. While the amount of transactions is similar between the two networks, Polygon has roughly three times more transfer events. \newline The output was stored in separate CSV files and also imported into a MongoDB\footnote{\url{https://www.mongodb.com/}} database. The database contains distinct collections for transactions and transfers. For additional information on the data fields of each collection, see Appendix.

@@

\subsection{Intra-set asset transfers}
In this thesis, we want to employ entity clustering heuristics to determine the number of entities represented by the remaining 137,544 addresses. Or, to phrase it differently, do users interact with Decentraland using multiple EOAs, and can we identify addresses that belong to the same user?
\newline
To achieve this, we often only need to consider transactions or token transfer events where both the ``\texttt{from}`` (sender) and the ``\texttt{to}`` (recipient) address are within the address set. We call this reduced dataset ``intra-set asset transfers``. Regarding normal transactions, this means that we only keep native asset transfers (Ether or Matic) as almost all addresses are EOAs. \newline
The intra-set asset transfer dataset is mainly used to generate the network graphs (see section xx). When clustering addresses using graph-based approaches, this has the advantage that we do not need to differentiate between EOA and smart contract account later on. Furthermore, it allows us to observe and visualize the asset flows between users. In total, we have 974,478 intra-set token transfers and 210,548 intra-set native asset transfers (101,621 Matic, 108,927 Ether).

\iffalse
%further simplifies the clustering process. \newline
%(logging in to Decentraland with multiple addresses) When it comes to entity clustering, we essentially want to answer the question of how many real-world entities our remaining  137,544 addresses represent. 
-Also for other clustering heuristics we often make this reduction, for example: In the self-approval approach, we only consider approval transactions where another address from the set is approved.
-filter our transactions and token transfers to entries where both the 'from' and the 'to' are within this narrowed pool of addresses. We define this as intra-set asset transfers. For the transactions, this means that we only consider native asset (Ether or matic) transfers. %between two EOAs
-As already mentioned in the introduction, the goal of this paper is to cluster addresses within who (1) were active in Decentraland during the 9-month period (e.g., within the initial dataset) and (2) transacted or interacted with any token on Ethereum or Polygon. 
-Essentialy, we want to answer two main research questions when it comes to clustering:
1. How many unique entities does the address set represent (i.e. were active on Decentraland during the 9-month period)?
2. Are users interacting through multiple accounts/EOAs/addresses with Decentraland?
-we focus on on-chain data (and not user location or avatar data), allowing us to exclude addresses that have not transacted or interacted with any token on Ethereum or Polygon. 
-As mentioned in the introduction, the goal of this paper is to cluster addresses belongig to the same entity that were recorded/active in Decentraland during the 9-month period. Therefore, we disregard addresses outside of this set. Also, since we only look at on-chain data, we cannot cluster addresses without activity on Polygon or Ethereum. Doing these two reductions, there are  141, 591 'clusterable' addresses remaining.
-Since all of our addresses are EOAs (some multisig smart contract wallets). -> Check this?
\fi

@@

Figure \ref{fig:ENS} illustrates the mechanism that allowed us to generate ground truth pairs in a simplified way: Initially, the domain name \texttt{`foo.eth`} resolves to address \texttt{`0xABC...`}. The owner has also set a reverse resolution for this address, indicated by the the edge from the address to the domain name. Next, the owner decides to change the resolved address to \texttt{`0xDEF...`}, again setting a reverse record.  As a domain can only point to one address at the time, the previous mapping from \texttt{`foo.eth`} to the old address is deleted. However, \texttt{`ABC....addr.reverse`} still points to \texttt{`foo.eth`} and when we query the name corresponding to \texttt{`0xABC...`} and \texttt{`0xDEF...`}, both will return \texttt{`foo.eth`}.
If this special case occurs, we assume that both addresses belong to the same entity.
%We did not observe cases where someone set a reverse record to a completely unrelated address (e.g. alice.eth)

@@

We applied the following methodology to find ground truth pairs: First, we gathered all addresses that interacted with ENS (NFTs) to reduce the amount of name look-ups. Second, we iterated over each remaining address and checked if it reverse-resolved to an ENS domain\footnote{This was done via the ENS module within web3.py \url{https://web3py.readthedocs.io/en/stable/ens_overview.html}, which provides an interface to look up domains and addresses.}. If so, we logged the ENS domain and address in a CSV file. Third, using our new list, we identified if two addresses referenced the same ENS domain. If matched, we recorded both addresses alongside the ENS domain. \newline
From this, we identified 11,440 reverse records and 40 address pairs, %(addresses pointing to the same .eth name)
which we utilized as our ground truth to evaluate various clustering heuristics. Although this sample is relatively small, it allows us to assess and draw conclusions about the effectiveness of the methods.%profiling techniques/methods
